---
title: "My BP Lab Overview (2018-06-29)"
author: "Brian M. Bot"
output: pdf_document
---

***

```{r global_options, include = FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE)
```

```{r loadLibraries, include=FALSE}
require(synapser)
require(tableone)
require(plyr)
require(dplyr)
require(utils)
require(knitr)
require(ggplot2)
synLogin()

```

```{r loadData, include=FALSE, cache=TRUE}

## SOME VARIABLES THAT ARE SET FOR USE THROUGHOUT
releasedVersions <- c("version 1.0.3, build 72",
                      "version 1.0.5, build 75",
                      "version 1.0.6, build 76",
                      "version 1.1.0, build 82")

## FUNCTION FOR PRINTING PRETTY TABLE
kableone <- function(x, ...) {
  capture.output(x <- print(x))
  knitr::kable(x, ...)
}

#####
## GET BACKGROUND SURVEY
#####
bg <- synTableQuery("SELECT * FROM syn11855399 WHERE uploadDate > '2018-03-14' AND uploadDate < '2018-06-30' AND dataGroups IS NULL AND userSharingScope='ALL_QUALIFIED_RESEARCHERS'")
aa <- as.data.frame(bg)
## USE ONLY RELEASED VERSIONS OF THE APP
aa <- aa[ which(aa$appVersion %in% releasedVersions), ]

## CENSOR AGE
aa$answers.age[ which(aa$answers.age > 89) ] <- 89

## CENSOR HEIGHT
## lower than 4 feet or higher than 7 foot six inches set as missing (very unlikely)
## lower than 5 feet or higher than 6 foot six inches censored at those values
aa$answers.height[ which(aa$answers.height < 48) ] <- NA
aa$answers.height[ which(aa$answers.height < 60) ] <- 60
aa$answers.height[ which(aa$answers.height > 90) ] <- NA
aa$answers.height[ which(aa$answers.height > 78) ] <- 78

## CENSOR WEIGHT
## lower than 80 or higher than 500 pounds set as missing (very unlikely)
## lower than 100 or higher than 300 pounds censored at those values
aa$answers.weight[ which(aa$answers.weight < 80) ] <- NA
aa$answers.weight[ which(aa$answers.weight < 100) ] <- 100
aa$answers.weight[ which(aa$answers.weight > 500) ] <- NA
aa$answers.weight[ which(aa$answers.weight > 300) ] <- 300

## KEEP MOST RECENT VERSION OF THE SURVEY
aa$createdOn <- as.POSIXct(aa$createdOn/1000, origin="1970-01-01")
aa <- aa[ order(aa$createdOn, decreasing = TRUE), ]
aa <- aa[ !duplicated(aa$healthCode), ]
aa <- aa[ order(aa$createdOn), ]

## GET DATA FROM BASELINE BP
bt <- synTableQuery("SELECT * FROM syn11860964 WHERE uploadDate > '2018-03-14' AND uploadDate < '2018-06-30' AND dataGroups IS NULL AND userSharingScope='ALL_QUALIFIED_RESEARCHERS'")
cc <- as.data.frame(bt)
## USE ONLY RELEASED VERSIONS OF THE APP
cc <- cc[ which(cc$appVersion %in% releasedVersions), ]
cc <- cc[ order(cc$createdOn, decreasing = TRUE), ]
cc <- cc[ !duplicated(cc$healthCode), ]
cc <- cc[ order(cc$createdOn), ]
cc <- cc[, c("healthCode", "answers.cuff")]
aa <- merge(aa, cc, by="healthCode", all.x=TRUE, all.y=FALSE)
aa$cuffCalibration <- aa$answers.cuff != "None"
names(aa) <- gsub(" ", "_", names(aa), fixed=TRUE)

#####
## LOOK AT USAGE PATTERNS
#####
uTab <- synTableQuery("SELECT * FROM syn11855393 WHERE uploadDate > '2018-03-14' AND uploadDate < '2018-06-30' AND dataGroups IS NULL AND userSharingScope='ALL_QUALIFIED_RESEARCHERS'")
uu <- as.data.frame(uTab)
uu <- uu[ which(uu$appVersion %in% releasedVersions), ]
uu$createdOn <- as.POSIXct(uu$createdOn/1000, origin="1970-01-01")
uu$createdOnDate <- as.Date(uu$createdOn)

## ADD TIME SINCE ENTERED STUDY
uu <- ddply(uu, .(healthCode), mutate,
            timeOn = createdOn - min(createdOn),
            daysOn = createdOnDate - min(createdOnDate))

## SUMMARIZE TO ONE PER HEALTHCODE
upp <- ddply(uu, .(healthCode), summarize,
             nBP = sum(originalTable == "Blood Pressure-v6"),
             nMorning = sum(originalTable %in% c("Morning-v4", "Morning-v5")),
             nMAB = sum(originalTable %in% c("Body and Mind-v6", "Body and Mind-v7")),
             nEvening = sum(originalTable %in% c("Night-v8", "Night-v9")),
             daysOnStudy = as.numeric(max(timeOn), units="days"))

## GET RIDE OF THEORETICAL DUPLICATE BP MEASUREMENTS
checkinTables <- c("Morning-v4", "Morning-v5", "Body and Mind-v6", "Body and Mind-v7", "Night-v8", "Night-v9")
allActivities <- uu[ !(uu$originalTable %in% c("Baseline-v2", checkinTables)), ]
allCheckins <- uu[ uu$originalTable %in% checkinTables, ]
allCheckins <- ddply(allCheckins, .(healthCode, createdOnDate), summarize,
                     morning = any(originalTable %in% c("Morning-v4", "Morning-v5")),
                     afternoon = any(originalTable %in% c("Body and Mind-v6", "Body and Mind-v7")),
                     evening = any(originalTable %in% c("Night-v8", "Night-v9")))
allCheckins <- ddply(allCheckins, .(healthCode), mutate,
                     daysOn = createdOnDate - min(createdOnDate))

```

```{r plottingFunctions, include=FALSE}

plotThreeWeeks <- function(df){
  tmp <- unique(df[, c("healthCode", "daysOn")])
  tmp <- tmp[ tmp$daysOn > 0 & tmp$daysOn < 22, ]
  tt <- as.data.frame(table(tmp$daysOn)/length(unique(tmp$healthCode)))
  twp <- ggplot(tt, aes(x=Var1, y=Freq)) +
    geom_histogram(stat="identity", fill="#2B688F") +
    xlab("days on study") +
    ylab("proportion") +
    theme_minimal()
  return(twp)
}

plotThreeWeeksCuff <- function(df){
  tmp <- unique(df[, c("healthCode", "daysOn", "cuffCalibration")])
  tmp <- tmp[ tmp$daysOn > 0 & tmp$daysOn < 22, ]
  tt1 <- as.data.frame(table(tmp$daysOn[which(tmp$cuffCalibration)])/length(unique(tmp$healthCode[which(tmp$cuffCalibration)])))
  tt1$cuff <- TRUE
  tt2 <- as.data.frame(table(tmp$daysOn[which(!tmp$cuffCalibration)])/length(unique(tmp$healthCode[which(!tmp$cuffCalibration)])))
  tt2$cuff <- FALSE
  tt <- rbind(tt1, tt2)
  twp <- ggplot(tt, aes(x=Var1, y=Freq, fill=cuff)) +
    geom_histogram(stat="identity", position="dodge") +
    scale_fill_manual(values=c("#5B95BA","#E566A1")) +
    xlab("days on study") +
    ylab("proportion") +
    theme_minimal()
  return(twp)
}

plotTimeOfDay <- function(df){
  tmp <- df
  ## APPROXIMATION OF LOCAL TIME
  tmp$hourOfDay <- floor(as.integer(format(tmp$createdOn, "%H")) + tmp$createdOnTimeZone/100)
  tmp$hourOfDay[ which(tmp$hourOfDay < 0) ] <- tmp$hourOfDay[ which(tmp$hourOfDay < 0) ] + 24
  tmp$hourOfDay[ which(tmp$hourOfDay >= 24) ] <- tmp$hourOfDay[ which(tmp$hourOfDay >= 24) ] - 24
  tt <- as.data.frame(table(tmp$hourOfDay)/nrow(tmp))
  tod <- ggplot(tt, aes(x=Var1, y=Freq)) +
    geom_histogram(stat="identity", fill="#2B688F") +
    xlab("local hour of the day") +
    ylab("proportion") +
    theme_minimal()
  return(tod)
}

plotDayOfWeek <- function(df){
  tmp <- df
  ## APPROXIMATION OF LOCAL TIME
  tmp$dayOfWeek <- factor(format(tmp$createdOn, "%a"), levels=c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"))
  tt <- as.data.frame(table(tmp$dayOfWeek)/nrow(tmp))
  dow <- ggplot(tt, aes(x=Var1, y=Freq)) +
    geom_histogram(stat="identity", fill="#2B688F") +
    xlab("day of the week") +
    ylab("proportion") +
    theme_minimal()
  return(dow)
}

```

## Table 1: Demographics

#### Overall

```{r tableOne, echo=FALSE}
#####
## DEMOGRAPHICS OVERVIEW
#####
catVars <- c("answers.sex", 
             "answers.ethnicity.Black_or_African_American", 
             "answers.ethnicity.Asian_or_Pacific_Islander",
             "answers.ethnicity.White",
             "answers.ethnicity.American_Indian_or_Alaska_Native",
             "answers.ethnicity.Other",
             "answers.education",
             "answers.tobacco",
             "answers.health")
contVars <- c("answers.age", "answers.height", "answers.weight")
tableOne <- CreateTableOne(vars=c(catVars, contVars), data=aa, factorVars=catVars)
kableone(tableOne, nonnormal=contVars)

```

\pagebreak

#### By cuff calibration

```{r tableOneByCuff, echo=FALSE}
tableOneStrata <- CreateTableOne(vars=c(catVars, contVars), data=aa, factorVars=catVars, strata="cuffCalibration")
kableone(tableOneStrata, nonnormal=contVars)
```

\pagebreak

## Usage Patterns

#### Time patterns for entering the study (demographics survey)

Slightly more individuals entered the study towards the end of the week or over the weekend.

```{r dowDemo, echo=FALSE, fig.height=4, fig.width=6}
print(plotDayOfWeek(aa))
```

Pattern for hour of the day when a user enters the study.

```{r todDemo, echo=FALSE, fig.height=4, fig.width=6}
print(plotTimeOfDay(aa))
```

\pagebreak

#### Time patterns for any BP measurement (check-in or on-demand)

There does not appear to be significant differences between days of the week.

```{r dowBP, echo=FALSE, fig.height=4, fig.width=6}
print(plotDayOfWeek(allActivities[allActivities$originalTable == "Blood Pressure-v6", ]))
```

Pattern for hour of the day when BP is measured.

```{r todBP, echo=FALSE, fig.height=4, fig.width=6}
print(plotTimeOfDay(allActivities[allActivities$originalTable == "Blood Pressure-v6", ]))
```

\pagebreak

#### 3 week retention

Decent retention in general (compared to other mHealth studies).

```{r tw, echo=FALSE, fig.height=4, fig.width=6}
print(plotThreeWeeks(allActivities))
```

Retention is MUCH better for participants who have a cuff calibration - especially at the long tail.

```{r twCuff, echo=FALSE, fig.height=4, fig.width=6}
tmpPlot <- merge(allActivities, aa[, c("healthCode", "cuffCalibration")], all.x=TRUE, all.y=FALSE)
print(plotThreeWeeksCuff(tmpPlot))
```

\pagebreak

#### 3 week check-in overview

```{r twci, echo=FALSE}
twci <- merge(allCheckins, aa[, c("healthCode", "cuffCalibration")], all.x=TRUE, all.y=FALSE)
twci <- twci[ twci$daysOn <= 21, ]
twci$m <- twci$morning & !twci$afternoon & !twci$evening
twci$a <- !twci$morning & twci$afternoon & !twci$evening
twci$e <- !twci$morning & !twci$afternoon & twci$evening
twci$ma <- twci$morning & twci$afternoon & !twci$evening
twci$me <- twci$morning & !twci$afternoon & twci$evening
twci$ae <- !twci$morning & twci$afternoon & twci$evening
twci$mae <- twci$morning & twci$afternoon & twci$evening

tmp <- twci[, c("m", "a", "e", "ma", "me", "ae", "mae")]
res <- colSums(tmp)
percent <- function(x, digits = 2, format = "f", ...) {
  paste0(formatC(100 * x, format = format, digits = digits, ...), " %")
}
```

Check-in | Percent of person-days
----- | -----
Morning only | `r as.integer(res[["m"]])` (`r percent(res[["m"]]/nrow(tmp))`)
Afternoon only | `r as.integer(res[["a"]])` (`r percent(res[["a"]]/nrow(tmp))`)
Evening only | `r as.integer(res[["e"]])` (`r percent(res[["e"]]/nrow(tmp))`)
Morning & Afternoon | `r as.integer(res[["ma"]])` (`r percent(res[["ma"]]/nrow(tmp))`)
Morning & Evening | `r as.integer(res[["me"]])` (`r percent(res[["me"]]/nrow(tmp))`)
Afternoon & Evening | `r as.integer(res[["ae"]])` (`r percent(res[["ae"]]/nrow(tmp))`)
Morning, Afternoon, & Evening | `r as.integer(res[["mae"]])` (`r percent(res[["mae"]]/nrow(tmp))`)

***

```{r twcic, echo=FALSE}
tmpc <- twci[ which(twci$cuffCalibration), c("m", "a", "e", "ma", "me", "ae", "mae")]
resc <- colSums(tmpc)
```

For only those with a cuff calibration:

Check-in | Percent of person-days
----- | -----
Morning only | `r as.integer(resc[["m"]])` (`r percent(resc[["m"]]/nrow(tmpc))`)
Afternoon only | `r as.integer(resc[["a"]])` (`r percent(resc[["a"]]/nrow(tmpc))`)
Evening only | `r as.integer(resc[["e"]])` (`r percent(resc[["e"]]/nrow(tmpc))`)
Morning & Afternoon | `r as.integer(resc[["ma"]])` (`r percent(resc[["ma"]]/nrow(tmpc))`)
Morning & Evening | `r as.integer(resc[["me"]])` (`r percent(resc[["me"]]/nrow(tmpc))`)
Afternoon & Evening | `r as.integer(resc[["ae"]])` (`r percent(resc[["ae"]]/nrow(tmpc))`)
Morning, Afternoon, & Evening | `r as.integer(resc[["mae"]])` (`r percent(resc[["mae"]]/nrow(tmpc))`)
